---
title: "The Cox model"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(survival)
library(survminer)
library(coxme)
library(coxphw)
library(splines) 
library(tidyverse) 
library(forcats)
library(car)
library(broom)
library(AICcmodavg)
library(sjPlot)

# library(sjPlot)
# library(sjmisc)
# library(sjlabelled) 
```

# Data

```{r}
eelife <- read.csv("../processed/eeAlldat.csv") %>%
  select(-X,cageID,larv_adult,larvalTreat,adultTreat,age,status,sex) 

# cols <- c("cageID", "larv_adult", "larvalTreat", "adultTreat","sex")
# eelife[cols] <- lapply(eelife[cols], factor)  
eelife$age <- as.numeric(eelife$age)
eelife$status <- as.numeric(eelife$status)
# sapply(eelife, class)

```

# A full model

Addressing non-proportional hazards: https://cran.r-project.org/web/packages/Greg/vignettes/timeSplitter.html
vignette("timedep", package = "survival")

```{r}
mod1.0 <- coxph(Surv(age, status) ~  larvalTreat + adultTreat + 
                   larvalTreat*adultTreat + sex + cluster(cageID), 
                 data = eelife)
summary(mod1.0) 
cox.zph(mod1.0) # very NPH

# extractAIC(mod1.0)
# tidy(mod1.0,exponentiate = TRUE)
# glance(mod1.0)

# Stratify on sex (cageID cannot be stratified)
mod1.1 <- coxph(Surv(age, status) ~  larvalTreat + adultTreat + 
                   larvalTreat*adultTreat + strata(sex) + 
                   cluster(cageID), data = eelife)
summary(mod1.1)
survfit(mod1.1) # estimates K different baseline hazard functions, one for each stratum
cox.zph(mod1.1) # PH
car::Anova(mod1.1, type="II")
plot(survfit(mod1.0, transform="km", global=TRUE))

# anova(mod1.0,mod1.1)
# Stratifying improves model, but still NPH

mod1.2 <- coxph(Surv(age, status) ~  larvalTreat + adultTreat + 
                   larvalTreat*adultTreat + larv_adult*strata(sex) +
                   cluster(cageID), data = eelife)
summary(mod1.2)
cox.zph(mod1.2) # NPH, due to interaction

mod1.2b <- coxph(Surv(age, status) ~ 
                   larv_adult*strata(sex), 
                   cluster=cageID, 
                data = eelife)
summary(mod1.2b)
cox.zph(mod1.2b)
#ggsurvplot(survfit(mod1.2b),eelife, conf.int = TRUE)
library(MuMIn)

mod1.3 <- coxph(Surv(age, status) ~  larvalTreat + strata(adultTreat) + 
                   larvalTreat*adultTreat + larv_adult*strata(sex), 
                data = eelife)
summary(mod1.3)
#autoplot(mod1.3)
anova(mod1.3)
#ggsurvplot(survfit(mod1.3),eelife, conf.int = TRUE)

mod1.4 <- coxph(Surv(age, status) ~  larvalTreat + strata(adultTreat) + 
                   larvalTreat*adultTreat, data = eelife)
summary(mod1.4)
cox.zph(mod1.4)
anova(mod1.4)
anova(mod1.3,mod1.4)
ggsurvplot(survfit(mod1.4),eelife, conf.int = TRUE)

mod1.5 <- coxph(Surv(age, status) ~  larvalTreat + adultTreat, data = eelife)
summary(mod1.5)
cox.zph(mod1.5)

mod1.6 <- coxph(Surv(age, status) ~  larvalTreat, data = eelife)
summary(mod1.6)
cox.zph(mod1.6)

mod1.7 <- coxph(Surv(age, status) ~  adultTreat, data = eelife)
summary(mod1.7)
cox.zph(mod1.7)

aictab(list(mod1.0,mod1.1,mod1.2,mod1.2b,mod1.3, mod1.4,mod1.5,
            mod1.6,mod1.7)) 

# mod1.3 is best (AICcWt=1.0)
# Although larv_adultDR_DR:strata(sex)M is (***), mod1.2 is bad (AICcWt=0.01)
# Even when larvalTreat*adultTreat is removed, mod1.2 is AICcWt=0.03
# We can conclude absence of sex-by-diet interaction, generally.

```

# Cage as random effect (mod1.2: larv_adultDR_DR:strata(sex)M is (***), but AICcWt=0.01)
# cluster(ID) argument for coxph = random effects argument for coxme
# i.e. model mod1.1 generally similar to mod1.2

```{r}
mod1.8 <- coxme(Surv(age, status) ~ larvalTreat + adultTreat + 
                  larvalTreat*adultTreat + strata(sex) +
                  (1 | cluster(cageID)), data = eelife)
summary(mod1.8) 
# Interaction not sig.
car::Anova(mod1.8, type="II")
# car::Anova(mod1.2, type="III"); use only if no interactions are involved

#anova(mod1.2b, mod1.3)
# cageID can be ignored
# Note: mod1.1 and mod1.3 are the same (i.e. cluster(cageID for coxph
# is same as (1 | cageID for coxme)))
# So, interpret mod1.1

#tab_model(mod1.0, mod1.1,mod1.2, mod1.3, file="AllDataMods.doc")

```

# For supplementary data
# Test for sex-by-diet interaction (individual life stage diet effects)
# Females

```{r}

eelifeF <- eelife %>%
  filter(sex=="F")

mod2.0 <- coxph(Surv(age, status) ~ larvalTreat + adultTreat + 
                     larvalTreat*adultTreat + cluster(cageID), 
                   data = eelifeF)
summary(mod2.0) 
cox.zph(mod2.0) # NPH
coef(mod2.0)

mod2.1 <- coxph(Surv(age, status) ~ larvalTreat + strata(adultTreat) + 
                  larvalTreat*adultTreat + cluster(cageID), data = eelifeF)
summary(mod2.1) 
cox.zph(mod2.1) # PH
# Looks good

mod2.2 <- coxph(Surv(age, status) ~ larvalTreat + strata(adultTreat) +
                 larvalTreat*adultTreat, data = eelifeF)
summary(mod2.2)
cox.zph(mod2.2) # PH

mod2.3 <- coxph(Surv(age, status) ~ larvalTreat + strata(adultTreat), 
                 data = eelifeF)
summary(mod2.3)
cox.zph(mod2.3) # PH

aictab(list(mod2.0, mod2.1, mod2.2, mod2.3))
# Model mod2.1 ties with mod2.2, and are both not so diff from mod2.3 
# Support to ignore cage
# Suggest the larv*adult interaction not so important

#tab_model(mod2.0, mod2.1, mod2.2,mod2.3, file="FemaleMods.doc")

```

# Males

```{r}
eelifeM <- eelife %>%
  filter(sex=="M")

mod3.0 <- coxph(Surv(age, status) ~ larvalTreat + adultTreat + 
                     larvalTreat*adultTreat + cluster(cageID), 
                   data = eelifeM)
summary(mod3.0) 
cox.zph(mod3.0) # NPH
coef(mod3.0)

mod3.1 <- coxph(Surv(age, status) ~ larvalTreat + strata(adultTreat) + 
                  larvalTreat*adultTreat + cluster(cageID), 
                 data = eelifeM)
summary(mod3.1) 
cox.zph(mod3.1) # NPH

mod3.2 <- coxph(Surv(age, status) ~ larvalTreat + strata(adultTreat) +
                 larvalTreat*adultTreat, data = eelifeM)
summary(mod3.2)
cox.zph(mod3.2) # NPH

mod3.3 <- coxph(Surv(age, status) ~ larvalTreat + strata(adultTreat), 
                 data = eelifeM)
summary(mod3.3)
cox.zph(mod3.3) # PH

aictab(list(mod3.0,mod3.1, mod3.2,mod3.3))
# Model mod2.1 ties with mod2.2: Support to ignore cage

tab_model(mod3.0, mod3.1, mod3.2,mod3.3, file="MaleMods.doc")

# The simplest model mod3.3 seems best (Wt 0.77)
# Mods 3.1 and 3.2 tie with Wt 0.11 - again suggesting cage can be ignored
# Larval-adult diet interaction not important for males
# Interpret the simplest mod

```

# Combined effect of larval-adult diets

```{r}
mod4.0 <- coxph(Surv(age, status) ~ larvalTreat + adultTreat + 
                  larv_adult + strata(sex) + cluster(cageID), 
                data = eelife)
summary(mod4.0)
cox.zph(mod4.0) # PH

mod4.1 <- coxph(Surv(age, status) ~ larvalTreat + adultTreat + 
                  larv_adult + strata(sex), 
                data = eelife)
summary(mod4.1)
cox.zph(mod4.1) # PH
anova(mod4.1)

mod4.2 <- coxph(Surv(age, status) ~ larvalTreat + adultTreat + 
                  larv_adult, data = eelife)
summary(mod4.2)
cox.zph(mod4.2) # PH

aictab(list(mod4.0, mod4.1, mod4.2))
# mod4.0 and mod4.1 are better, but tie. Seems cage can be ignored

# tab_model(mod1.1,mod1.2,mod4.0,mod4.1,mod4.2, file="AllDataMods.doc")

```

# Combined effect of sex-by-larval_adult diet effects
# There are sig S-by-L diet effects - diet affects female lifespan

```{r}
# Females (effects)

mod5.0 <- coxph(Surv(age, status) ~ larvalTreat + strata(adultTreat) + 
                     larv_adult + cluster(cageID), data = eelifeF)
summary(mod5.0) # cage effects in 1 cage
cox.zph(mod5.0) # PH

mod5.1 <- coxph(Surv(age, status) ~ larvalTreat + strata(adultTreat) + 
                     larv_adult, data = eelifeF)
summary(mod5.1) 
cox.zph(mod5.1) # PH

mod5.2 <- coxph(Surv(age, status) ~ adultTreat + larv_adult, 
                   data = eelifeF)
summary(mod5.2) 
cox.zph(mod5.2) # NPH

mod5.3 <- coxph(Surv(age, status) ~ larv_adult, 
                   data = eelifeF)
summary(mod5.3) 
cox.zph(mod5.3) # PH

aictab(list(mod5.0,mod5.1, mod5.2, mod5.3))
anova(mod5.1, mod5.2)
# mod1.2.1 is best


# Males # No effects
mod6.0 <- coxph(Surv(age, status) ~ larvalTreat + strata(adultTreat) + 
                     larv_adult + cluster(cageID), data = eelifeM)
summary(mod6.0) # cage effects in 1 cage
cox.zph(mod6.0) # PH

mod6.1 <- coxph(Surv(age, status) ~ larvalTreat + strata(adultTreat) + 
                     larv_adult, data = eelifeM)
summary(mod6.1) 
cox.zph(mod6.1) # NPH

mod6.2 <- coxph(Surv(age, status) ~ adultTreat + larv_adult, 
                   data = eelifeM)
summary(mod6.2) 
cox.zph(mod6.2) # PH

mod6.3 <- coxph(Surv(age, status) ~ larv_adult, 
                   data = eelifeM)
summary(mod6.3) 
cox.zph(mod6.3) # NPH

aictab(list(mod6.0, mod6.1, mod6.2, mod6.3))

anova(mod6.1, mod6.2, mod6.3)

```

# Final models

```{r}
mf1 <- coxph(Surv(age, status) ~ larv_adult * strata(sex) + cageID, eelife)
summary(mf1)
cox.zph(mf1)

mf2 <- coxph(Surv(age, status) ~ larv_adult * strata(sex), eelife)
summary(mf2)
cox.zph(mf2)

mf3 <- coxph(Surv(age, status) ~ larv_adult + cageID, eelife)
summary(mf3)
cox.zph(mf3)

mf4 <- coxph(Surv(age, status) ~ larv_adult + strata(sex), eelife)
summary(mf4)
cox.zph(mf4)

mf5 <- coxph(Surv(age, status) ~ larv_adult + strata(sex) + cageID, eelife)
summary(mf5)
cox.zph(mf5)

mf6 <- coxph(Surv(age, status) ~ larv_adult, eelife)
summary(mf6)
cox.zph(mf6)

aictab(list(mf1,mf2,mf3,mf4,mf5,mf6))
```

