---
title: "The Cox model"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(survival)
library(survminer)
library(splines) #needed by survival package
library(tidyverse) 
library(forcats)
library(cowplot)
library(Greg)

```

# CoxPH models 

Assupmtion: Proportional hazards (effects of all covariates do not change over time)
Non-PH therefore under- or overestimated risk

PH assumption (i.e. residuals independent of time) violated
Early: Adult diet interaction sig, albeit skewed residuals
Interaction of non-PH variables with age non-convergent (diet interacts heavily with age - nothing else is significant)
Models on age-classes not significant for either diet, but a large effect of sex.

Larval treatment effect depends on adult age 


# All terms and a 3-way interaction

Interaction is not significant
Non-PH in larval diet and sex
see Fox and Weisberg (2018)

```{r}
eelife <- read.csv("../processed/eeAlldat.csv") %>%
  select(-X,cageID,larv_adult,larvalTreat,adultTreat,age,status,sex)

cph1 <- coxph(Surv(age, status) ~ larvalTreat + adultTreat + sex +
                larvalTreat:adultTreat:sex, 
              data = eelife)
summary(cph1)
#ggforest(cph1, data = eelife)
test.cph1 <- cox.zph(cph1)
test.cph1
#ggcoxzph(test.cph1)

```

# Code interaction as a compound variable (2 diets and sex)

All interactions are significant - likely due to sex
cpd_var is non-PH

```{r}
eelife <- eelife %>% unite(cpd_var, larv_adult, sex, 
                           sep = "_", remove = FALSE)

cph2 <- coxph(Surv(age, status) ~ cpd_var, data = eelife)
summary(cph2)
cox.zph(cph2)

```

# Fit sex additively in cph2

All terms are significant
Sex is non-PH

```{r}
cph3 <- coxph(Surv(age, status) ~ larv_adult + sex, data = eelife)
summary(cph3)
cox.zph(cph3)
```

# Fit larval diet additively

Terms with sex are all sig, larval diet nsig
Both covariates non-PH

```{r}
eelife <- eelife %>% unite(adTreat_sex, adultTreat, sex, 
                           sep = "_", remove = FALSE)

cph4 <- coxph(Surv(age, status) ~ larvalTreat + adTreat_sex, 
              data = eelife)
summary(cph4)
#ggforest(cph2, data = eelife)
cox.zph(cph4)
```

# Take from simple models
1. LOOKS LIKE SEX HAS A MAJOR IMPACT - CONSIDER SEXES SEPARATELY
2. DEAL WITH NON-PH COVARIATES





# 2. Stratify the time variable
```{r}
hist(eelife$age)

eelife$age.intv <- car::recode(eelife$age, 
                              "lo:19=1; 
                              20:39=2; 
                              40:59=3; 
                              60:79=4; 
                              80:hi=5") 
xtabs(~ age.intv, data=eelife)

cph4 <- coxph(Surv(age, status) ~ larvalTreat + adultTreat + sex + 
                    strata(age.intv), data=eelife)
cph4

test.cph4 <- cox.zph(cph4)
test.cph4
#ggcoxzph(test.cph4)
# Sex and global p-vals are still non-proportional


# Add interaction back
# cph5 <- coxph(Surv(age, status) ~ adultTreat + larvalTreat + sex + 
#                     larvalTreat:adultTreat,strata(age.intv),
#               data=eelife)
# cph5
# ‘floor’ not meaningful for factors


```

# Cox model with split-time variable

Unsplit-time variable: Adult treatment has a large effect. Sex has a large effect. Larval treat and sex have very non-PH

Split-time variable: Adult treatment and sex have large effects. Larval treat and sex have non-PH

Splitting at interval of 0.5: Both treatments have PHs, but sex is still very much confounding.

```{r}

eelife <- eelife %<>% 
  mutate(status = factor(status),
        larvalTreat = factor(larvalTreat),
        adultTreat = factor(adultTreat),
        sex = factor(sex))



# A regular cox regression
## Requires a defined status level

regular_model <- coxph(Surv(age, status==1) ~ larvalTreat + 
                         adultTreat + sex, data = eelife)
summary(regular_model)
cox.zph(regular_model)


# Now, split dataset and repeat model

spl_eelife <-
  eelife %>% 
  timeSplitter(by = 0.5,
               event_var = "status",
               event_start_status = 0,
               time_var = "age")

interval_model <-
  update(regular_model, 
         Surv(Start_time, Stop_time, status == 1) ~ .,
         data = spl_eelife)

summary(interval_model)

# The difference between the models should be negligible:

cbind(Regular = coef(regular_model),
      Interval = coef(interval_model),
      Difference = coef(regular_model) - coef(interval_model)) %>% 
  txtRound(digits = 5) %>% 
  knitr::kable(align = "r")


# Now, look for time varying coefficients with survival::cox.zph() 

cox.zph(regular_model) %>% 
  extract2("table") %>% 
  txtRound(digits = 2) %>% 
  knitr::kable(align = "r")

# larvalTreat and sex are time-variable


# Update split-time model with interaction

time_int_model <- 
  update(interval_model,
         .~.+larvalTreat:Start_time,
         .~.+sex:Start_time)
summary(time_int_model)

sx <- cox.zph(time_int_model)
sx

ggcoxzph(sx)

#ggforest(time_int_model, data = spl_eelife)

ggforest(
  time_int_model,
  data = spl_eelife,
  main = NULL,
  cpositions = c(0.02, 0.22, 0.4),
  fontsize = 0.7,
  refLabel = "reference",
  noDigits = 2
)



# Adjusted survival curve for cox model
# ggadjustedcurves(regular_model, data = eelife,
#                     variable  = eelife[, "larvalTreat"],
#                    legend.title = "Larval Treatment", )
```



## Sexes separately (to remove confounding effects of sex)

PH assumption violated
Intearction of non-PH variables with age non-convergent (adultTrt sig)
Models on age-classes not significant for either diet


### Females

Unsplit model: Adult treat highly sig; but adult treat has non-PH

Split model: At 0.5, adult treat remains highly sig; Adult treatment depends on age very significantly; All hazards are proportional!

This looks like a valid solution.

```{r}
female <- subset(eelife, eelife$sex=="F")

female <- female %>% 
  mutate(status = factor(status),
        larvalTreat = factor(larvalTreat),
        adultTreat = factor(adultTreat),
        sex = factor(sex))


# Fit a regular cox model
regular_model <- coxph(Surv(age, status == 1) ~ larvalTreat + 
                         adultTreat, data = female)
summary(regular_model)
cox.zph(regular_model)


# Now, split dataset and repeat model

spl_female <-
  female %>% 
  timeSplitter(by = 0.5,
               event_var = "status",
               event_start_status = 0,
               time_var = "age")

interval_model <-
  update(regular_model, 
         Surv(Start_time, Stop_time, status == 1) ~ .,
         data = spl_female)

summary(interval_model)


# The difference between the models should be negligible:

cbind(Regular = coef(regular_model),
      Interval = coef(interval_model),
      Difference = coef(regular_model) - coef(interval_model)) %>% 
  txtRound(digits = 5) %>% 
  knitr::kable(align = "r")


# Now, look for time varying coefficients with survival::cox.zph() 

cox.zph(regular_model) %>% 
  extract2("table") %>% 
  txtRound(digits = 2) %>% 
  knitr::kable(align = "r")

# adultTreat is time-variable


# Update split-time model with interaction

time_int_model <- 
  update(interval_model,
         .~.+adultTreat:Start_time)
summary(time_int_model)

sx <- cox.zph(time_int_model)
sx

#ggcoxzph(sx)

#ggforest(time_int_model, data = spl_eelife)

FF <- ggforest(
  time_int_model,
  data = spl_female,
  main = NULL,
  cpositions = c(0.02, 0.22, 0.4),
  fontsize = 0.7,
  refLabel = "reference",
  noDigits = 2
)

FF

# ggsave(FF,filename = "~/MyGithub/early_experience/Plots/Female_Greg_Forest_HR.pdf", height=3, width=6)

```

### Males

Unsplit model: Adult DR is higly sig in males also; Both larval and adult treat are non-PH.

Split model: Has larval treat sig also (P=0.0195), with PHs for both larval treat and larval treat*time; However, non-PH for adult treat.

This seems a satisfactory solution since we are interested in larval diet.

```{r}

male <- subset(eelife, eelife$sex=="M")

male <- male %>% 
  mutate(status = factor(status),
        larvalTreat = factor(larvalTreat),
        adultTreat = factor(adultTreat),
        sex = factor(sex))


# Fit a regular cox model
regular_model <- coxph(Surv(age, status==1) ~ larvalTreat + 
                         adultTreat, data = male)
summary(regular_model)
cox.zph(regular_model)


# Now, split dataset and repeat model

spl_male <-
  male %>% 
  timeSplitter(by = 0.5,
               event_var = "status",
               event_start_status = 0,
               time_var = "age")

interval_model <-
  update(regular_model, 
         Surv(Start_time, Stop_time, status == 1) ~ .,
         data = spl_male)

summary(interval_model)


# The difference between the models should be negligible:

cbind(Regular = coef(regular_model),
      Interval = coef(interval_model),
      Difference = coef(regular_model) - coef(interval_model)) %>% 
  txtRound(digits = 5) %>% 
  knitr::kable(align = "r")


# Now, look for time varying coefficients with survival::cox.zph() 

cox.zph(regular_model) %>% 
  extract2("table") %>% 
  txtRound(digits = 2) %>% 
  knitr::kable(align = "r")

# both variables are time-nonrandom


# Update split-time model with interaction

time_int_model <- 
  update(interval_model,
         .~.+larvalTreat:Start_time,
         .~.+adultTreat:Start_time)
summary(time_int_model)

sx <- cox.zph(time_int_model)
sx
#adultTreat still non-PH

#ggcoxzph(sx)

#ggforest(time_int_model, data = spl_eelife)

MM <- ggforest(
  time_int_model,
  data = spl_male,
  main = NULL,
  cpositions = c(0.02, 0.22, 0.4),
  fontsize = 0.7,
  refLabel = "reference",
  noDigits = 2
)
MM

ggsave(FF,filename = "~/MyGithub/early_experience/Plots/male_Greg_Forest_HR.pdf", height=3, width=6)
```

# Diagnostics

```{r}
## Influential cases

# typeof residuals on Y axis: c(“martingale”, “deviance”, “score”, “schoenfeld”, “dfbeta”, “dfbetas”, “scaledsch”, “partial”).
# lin_mm <- ggcoxdiagnostics(time_int_model, type = "dfbeta",
#                  linear.predictions = FALSE, ggtheme = theme_bw())
# 
# ggsave(lin_mm,filename = "~/MyGithub/early_experience/Plots/male_nonlinearity_test.pdf", height=5, width=9)


### Visualize deviance residuals to check for outliers

# dev_mm <- ggcoxdiagnostics(time_int_model, type = "deviance",
#                  linear.predictions = FALSE, ggtheme = theme_bw())
# 
# dev_reg <- ggcoxdiagnostics(interval_model, type = "deviance",
#                  linear.predictions = FALSE, ggtheme = theme_bw())
# 
# ggsave(dev_mm,filename = "~/MyGithub/early_experience/Plots/male_deviance_test.pdf", height=5, width=9)


## Testing non-linearity
### i.e. checking if functional form of model is correcctly specified

# ggcoxfunctional(Surv(age, status==2) ~ age + log(age) + sqrt(age), data = male)
```

