---
title: "Early Experience Factorial Experiment"
author: "Enoch Ng'oma"
date: "7/4/2018"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

# Analysis of diet-dependent developmental plasticity in D. melanogaster lifespan

# Load packages

```{r}
library(survival)
library(survminer)
library(splines) #needed by survival package
library(tidyverse) 
library(forcats)
library(cowplot)
library(Greg)


source("color_map.R")
source("ggplot_theme.R")
```

## Load data & combine female and male events

```{r}
# Female data
eelifeF <- read.table('../processed/Female_events_eelife.txt',
                     sep = "\t", header = TRUE,
                     stringsAsFactors = FALSE) 

eelifeF <- select(eelifeF, -setDate, -flipDate, -age)
names(eelifeF)[1:2]<-c("age", "cageID")
eelifeF$sex <- "F"

# Male data
eelifeM <- read.table('../processed/Male_events_eelife.txt',
                     sep = "\t", header = TRUE,
                     stringsAsFactors = FALSE) 

eelifeM <- select(eelifeM, -setDate, -flipDate, -age)
names(eelifeM)[1:2]<-c("age", "cageID")
eelifeM$sex <- "M"
  
eelife <- rbind(eelifeF,eelifeM)
  

#eelife$carried <- 0 
eelife <- na.omit(eelife)

eelife$status <- ifelse(eelife[,"status"] == 2, 1, 0)

#write.csv(eelife, file="../processed/eeAlldat.csv, row.names = FALSE")
```

# Descriptives

```{r, dens}
cohort <- c('F&M','FF','MM')

dens.list <- vector(mode='list', length=3)
names(dens.list) <- cohort

for (ii in cohort) {
  
if(ii=='F&M')
{
# Male and female together

pp <- ggplot(eelife, aes(age, color = larv_adult)) + 
    geom_density(alpha = 0.5) +
    coord_cartesian(xlim = c(-10, 120), ylim = c(0, 0.03)) + 
    #adjust frame of x-axis
  
    scale_x_continuous(expand = c(0, 0), limits = c(-10, 120)) +
    expand_limits(x=125) +
    #limit plot area on x-axis
  
    scale_y_continuous(expand = c(0, 0)) +
    tx_color_map() +
    theme(axis.title = element_text(size = 12)) +
    xlab("Age (days)") +
    ylab("Density") +
    theme_half_open() +
    theme(legend.position = "none") 

    #element.text = 12
    #tx_fill_map() +
    #my_theme
#fm
}
  
if(ii=='FF')
  
{

# Female alone

pp <- ggplot(eelifeF, aes(age, color = larv_adult)) + 
    geom_density(alpha = 0.5) +
    coord_cartesian(xlim = c(-10, 120), ylim = c(0, 0.03)) + 
    #adjust frame of x-axis
    scale_x_continuous(expand = c(0, 0), limits = c(-10, 120)) +
    expand_limits(x=125) +
    #limit plot area on x-axis
    scale_y_continuous(expand = c(0, 0)) +
    tx_color_map() +
    theme(axis.title = element_text(size = 12)) +
    xlab("Age (days)") +
    ylab("Density") +
    theme_half_open() +
    theme(legend.position = "none") 
#ff

}
  
if(ii=='MM')

{
# Male alone

pp <- ggplot(eelifeM, aes(age, color = larv_adult)) + 
    geom_density(alpha = 0.5) +
    coord_cartesian(xlim = c(-10, 120), ylim = c(0, 0.03)) + 
    #adjust frame of x-axis
    scale_x_continuous(expand = c(0, 0), limits = c(-10, 120)) +
    expand_limits(x=125) +
    #limit plot area on x-axis
    scale_y_continuous(expand = c(0, 0)) +
    tx_color_map() +
    theme(axis.title = element_text(size = 12)) +
    xlab("Age (days)") +
    ylab("Density") +
    theme_half_open() +
    theme(legend.position = c(0.7, 0.8)) 
#mm
}
dens.list[[ii]] <- pp
}

densPlot <- plot_grid(dens.list[[1]],dens.list[[2]],dens.list[[3]],
                      nrow =3, ncol=1, 
                      labels = c('a.', 'b.', 'c.'), 
                      align = 'h', 
                      rel_widths = c(1.5,1.5, 1.5), 
                      label_size = 10)
densPlot
# ggsave(last_plot(), file = "../plots/density_male_female.pdf",
#        width = 6, height = 4)
```

# Testing developmental theories

## a) Survival
#### Males and females together
#### Females alone
#### Males alone
## b) Reaction norms


# Male and female combined

```{r}

eelife <- read.csv("../processed/eeAlldat.csv") %>%
  select(-X)

# Look at distribution of events
eelife <- eelife %>% 
  mutate(stats = ifelse(status == 2, 1, 0)) %>%
  select(cageID,larvalTreat,adultTreat,age,status,sex)


mf <- survfit(Surv(age, status) ~ larvalTreat + adultTreat + sex,
                   conf.type = "log", 
                   conf.int = 0.95,
                   type = "kaplan-meier",
                   error = "greenwood",
                   data = eelife)

summary(mf)$table[, "median"]
sum_survfit <- surv_summary(mf, data = eelife)

write.csv(sum_survfit, 
          file = "../processed/SummarySurvModel_larvAdulSex.csv",
          row.names = FALSE) 

# Write quantiles
write.csv(quantile(mf, probs = c(0.25, 0.5, 0.75)), 
          file = "../processed/quantiles.csv")


mf.plot <- ggsurvplot(mf, data = eelife, 
                     conf.int = FALSE,
                     break.time.by = 20,
                     palette = pal,
                     fun = "pct", 
                     surv.median.line = "hv",
                    # font.y = c(14),
                     font.x = c(font_size),
                     font.y = c(font_size),
                     font.tickslab = c(font_size),
                     #font.main = 11,
                     #risk.table = FALSE,
                     legend = "none",
                     risk.table = TRUE,
                     risk.table.col = "strata",
                     risk.table.height = 0.5,
                    # title = "Females & Males",
                     legend.labs = c("C-C-F","C-C-M",
                                     "C-DR-F","C-DR-M",
                                     "DR-C-F","DR-C-M",
                                     "DR-DR-F","DR-DR-M"),
                     legend.title = "Diet-Sex \n combination") +
                     guides(colour = guide_legend(nrow = 8)) +
                     theme_survminer(font.main = 11)  +
  xlab("Age (days)")
mf.plot

ggsave(filename = "~/MyGithub/early_experience/plots/SurvPlots_for_fullModel_with_RiskTable.pdf", height=6, width=8, print(mf.plot))

# Test significance
survdiff(Surv(age) ~larvalTreat+adultTreat+sex,
                     data = eelife, rho = 0) 

#quantile(mf,  probs = c(0.1, 0.5, 0.9))
```

# EarlyRich (C diet) 

### Females only

```{r}
earlyRichF <- subset(eelifeF, eelifeF$larvalTreat=="C")

erf <- survfit(Surv(age, status == 2) ~ adultTreat,
                   conf.type = "log", 
                   conf.int = 0.95,
                   type = "kaplan-meier",
                   error = "greenwood",
                   data = earlyRichF)

summary(erf)$table[, "median"]

erf.plot <- ggsurvplot(erf, data = earlyRichF, 
                     conf.int = FALSE,
                     break.time.by = 20,
                     palette = c("#999999", "#E69F00", 
                                 "#56B4E9", "#009E73"),
                     fun = "pct", 
                     surv.median.line = "hv",
                     font.x = c(font_size),
                     font.y = c(font_size),
                     font.tickslab = c(font_size),
                     #font.main = 12,
                     risk.table = FALSE,
                     #title="A",  
                     #legend="none", #does not work
                     legend = c(0.9, 0.8),                     
                     title = "Early-C Females",
                     legend.labs = c("C", "DR"),
                     legend.title = "Adult Diet") +
  theme_survminer(font.main = 11)  +
  xlab("Age (days)")
erf.plot

# Test significance
survdiff(Surv(age, status==2) ~ adultTreat,
                     data = earlyRichF, rho = 0) 

#quantile(f_er1,  probs = c(0.1, 0.5, 0.9))

```

# EarlyRich (C diet) 

### Males only

```{r}
earlyRichM <- subset(eelifeM, eelifeM$larvalTreat=="C")

erm <- survfit(Surv(age, status == 2) ~ adultTreat,
                   conf.type = "log", 
                   conf.int = 0.95,
                   type = "kaplan-meier",
                   error = "greenwood",
                   data = earlyRichM)

summary(erm)$table[, "median"]

erm.plot <- ggsurvplot(erm, data = earlyRichM, 
                     conf.int = FALSE,
                     break.time.by = 20,
                     palette = c("#999999", "#E69F00", 
                                 "#56B4E9", "#009E73"),
                     fun = "pct", 
                     surv.median.line = "hv",
                     font.x = c(font_size),
                     font.y = c(font_size),
                     font.tickslab = c(font_size),
                     #font.main = 12,
                     risk.table = FALSE,
                     #legend="none",
                     legend = c(0.9, 0.8),                     
                     title = "Early-C Males",
                     legend.labs = c("C", "DR"),
                     legend.title = "Adult Diet") +
  theme_survminer(font.main = 11) +
  xlab("Age (days)")
erm.plot

survdiff(Surv(age) ~ adultTreat,
                     data = earlyRichM, rho = 0) 

```

# EarlyPoor (DR diet) 

### Females only

```{r}
earlyPoorF <- subset(eelifeF, eelifeF$larvalTreat=="DR")

epf <- survfit(Surv(age, status==2) ~ adultTreat,
                   conf.type = "log", 
                   conf.int = 0.95,
                   type = "kaplan-meier",
                   error = "greenwood",
                   data = earlyPoorF,)
summary(epf)$table[, "median"]

epf.plot <- ggsurvplot(epf, data = earlyPoorF, 
                     conf.int = FALSE,
                     break.time.by = 20,
                     palette = c("#999999", "#E69F00", 
                                 "#56B4E9", "#009E73"),
                     fun = "pct", 
                     surv.median.line = "hv",
                     font.x = c(font_size),
                     font.y = c(font_size),
                     font.tickslab = c(font_size),
                     #font.main = 12,
                     risk.table = FALSE,
                     #legend="none",
                     legend = c(0.9, 0.8),                     
                     title = "Early-DR Females",
                     legend.labs = c("C", "DR"),
                     legend.title = "Adult Diet") +
  theme_survminer(font.main = 11)  +
  xlab("Age (days)")
epf.plot

# Test significance
survdiff(Surv(age, status==2) ~ adultTreat,
                     data = earlyPoorF, rho = 0) 

#quantile(f_er1,  probs = c(0.1, 0.5, 0.9))

```

### Males only

```{r}
earlyPoorM <- subset(eelifeM, eelifeM$larvalTreat=="DR")

epm <- survfit(Surv(age, status == 2) ~ adultTreat,
                   conf.type = "log", 
                   conf.int = 0.95,
                   type = "kaplan-meier",
                   error = "greenwood",
                   data = earlyPoorM)

summary(epm)$table[, "median"]

epm.plot <- ggsurvplot(epm, data = earlyPoorM, 
                     conf.int = FALSE,
                     break.time.by = 20,
                     palette = c("#999999", "#E69F00", 
                                 "#56B4E9", "#009E73"),
                     fun = "pct", 
                     surv.median.line = "hv",
                     font.x = c(font_size),
                     font.y = c(font_size),
                     font.tickslab = c(font_size),
                     #font.main = 12,
                     risk.table = FALSE,
                     #legend="none",
                     legend = c(0.9, 0.8),                     
                     title = "Early-DR Males",
                     legend.labs = c("C", "DR"),
                     legend.title = "Adult Diet") +
  theme_survminer(font.main = 11) +
  xlab("Age (days)")
epm.plot

survdiff(Surv(age, status ==2) ~ adultTreat,
                     data = earlyPoorM, rho = 0) 
```

# Arrange plots

```{r}
splots <- list()
splots[[1]] <- erf.plot
splots[[2]] <- erm.plot
splots[[3]] <- epf.plot
splots[[4]] <- epm.plot


if (FALSE) {
# Arrange and save into pdf file
res <- arrange_ggsurvplots(splots,ncol = 2, nrow = 2,
                            print = FALSE)
}

ggsave(filename = "~/MyGithub/early_experience/plots/SurvPlots_for_larvalDiet.pdf", height=6, width=8,res)

```





