---
title: "Mixed effects Cox & Bayesian Regression"
author: "Enoch Ng'oma"
date: "7/4/2018"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
editor_options: 
  chunk_output_type: console
---

# Analysis of diet-dependent developmental plasticity in D. melanogaster lifespan

# Load packages

```{r}
library(survival)
#library(coxme)
library(rstanarm)
library(tidyverse)
library(cowplot)
library(tidybayes)
library(bayesplot)
library(broom)

#source("color_map.R")
#source("ggplot_theme.R")
```

## Load data & combine female and male events

```{r}

# Data has both male and female events

eelife <- read.csv("../processed/eeAlldat.csv") %>%
  select(-X)

```

# Basic model

age ~ larv + adult +sex + cluster(larv) : LarvalTreatDR becomes very sig

```{r}

M <- coxph(Surv(age, status==2) ~ larvalTreat + adultTreat + sex, 
                   data = eelife)
M

# Model with robust SE via clustering
M2 <- coxph(Surv(age, status==2) ~ larvalTreat + 
              adultTreat + sex + cluster(larvalTreat), data = eelife)
M2
``` 

# Mixed effects model

```{r}
M3 <- coxme(Surv(age, status==2) ~ larvalTreat + adultTreat + 
              sex  + (1 | cageID), data = eelife)
M3

M4 <- coxme(Surv(age, status==2) ~ larvalTreat + sex  + 
              (1 | adultTreat), data = eelife)
M4

M5 <- coxme(Surv(age, status==2) ~ larvalTreat + adultTreat + sex +
              (1 | cageID), data = eelife)
M5

M6 <- coxme(Surv(age, status==2) ~ larvalTreat + adultTreat +
              (1 | sex), data = eelife)
M6

M7 <- coxme(Surv(age, status==2) ~ larvalTreat + adultTreat +
              (1 | sex) + (1 | cageID), data = eelife)
M7
```


# Bayesian survival models

# rstanarm

## Non-PHs modelled with a smooth cubic B-spline function for the time-varying log hazard ratio

```{r}

eelife <- read.csv("../processed/eeAlldat.csv") %>%
  select(-X)

start_time <- Sys.time()

set.seed(989898)
test_CHAINS <- 4
#test_CORES  <- 1
test_SEED   <- 12345
test_ITER   <- 1000

### M-splines is the default in stan_surv() 

# Baseline hazard function - assuming covariate effect of zero
mod1 <- stan_surv(formula = Surv(age, status) ~ tve(larvalTreat) + 
                     tve(adultTreat) + tve(sex), 
                  data    = eelife,
                  chains  = 8, 
                  cores   = 4, 
                  seed    = 414867,
                  iter    = 1e5)
print(mod1, digits = 3)
plot(mod1) + coord_cartesian(ylim = c(0, 0.1))
# 1000 iter for 30 min

# Test different parametric hazards
# i.e. fit several models, each with a different baseline hazard

mod1_exp      <- update(mod0, basehaz = "exp") #47min
mod1_weibull  <- update(mod0, basehaz = "weibull") # ~ 2hrs
mod1_gompertz <- update(mod0, basehaz = "gompertz") # 94s - not converged
mod1_bspline  <- update(mod0, basehaz = "bs") # 98.5 min
mod1_mspline1 <- update(mod0, basehaz = "ms") # 52.9 min
mod1_mspline2 <- update(mod0, basehaz = "ms", 
                        basehaz_ops = list(df = 10))

# save(mod0,mod1_exp,mod1_weibull,mod1_gompertz,mod1_bspline,
#        mod1_mspline1,mod1_mspline2, file="../processed/BaseHaz.Rda")


# Plot the baseline hazards with 95% posterior uncertainty limits
load("../processed/BaseHaz.Rda")

plotfun <- function(model, title) {
  plot(model, plotfun = "basehaz") +              
    coord_cartesian(ylim = c(0,0.4)) +            
    labs(title = title) +                        
    theme(plot.title = element_text(hjust = 0.5,size = 11),
          axis.title = element_text(size = 10)) + 
    theme_set(theme_cowplot()) +
    theme_half_open() 
}
p_exp      <- plotfun(mod1_exp,      title = "Exponential")
p_weibull  <- plotfun(mod1_weibull,  title = "Weibull")
p_gompertz <- plotfun(mod1_gompertz, title = "Gompertz")
p_bspline  <- plotfun(mod1_bspline,  title = "B-splines with df = 5")
p_mspline1 <- plotfun(mod1_mspline1, title = "M-splines with df = 5")
p_mspline2 <- plotfun(mod1_mspline2, title = "M-splines with df = 10")
cs <-bayesplot::bayesplot_grid(p_exp,
                          p_weibull,
                          p_gompertz,
                          p_bspline,
                          p_mspline1,
                          p_mspline2,
                          grid_args = list(ncol = 3))

ggplot2::ggsave(cs,filename="../plots/baseHaz_tve_fixedEffs.pdf",height = 6, width=10)

# Compare fits using `loo` for `stansurv` objects

lcomp <- loo_compare(loo(mod1_exp),
            loo(mod1_weibull),
            loo(mod1_gompertz),
            loo(mod1_bspline),
            loo(mod1_mspline1),
            loo(mod1_mspline2))

## Record execution time and multicore use
end_time <- Sys.time()
diff_time <- end_time - start_time


```

# Non-PHs modelled with a piecewise constant function

```{r}
# Piecewise constant: alternative to model tve() log hazard ratio

mod2 <- stan_surv(formula = Surv(age, status==2) ~ 
                    tve(larvalTreat, degree = 0, knots = 2.5) + 
                    tve(adultTreat, degree = 0, knots = 2.5) + 
                    tve(sex, degree = 0, knots = 2.5),
                    data    = eelife,
                    basehaz = "ms", basehaz_ops = list(df = 10),
                    chains  = test_CHAINS, 
                    cores   = test_CORES, 
                    seed    = test_SEED,
                    iter    = test_ITER)

#save(mod2, file="../processed/pieceW_tve.Rda")
load("../processed/pieceW_tve.Rda")

plot(mod2) + coord_cartesian(ylim = c(0, 0.1)) # basehaz
summary(mod2)

plot(mod2, pars = "tve")

```


# Multilevel models

```{r}
seed <- set.seed(787878)
chains <- 12
cores  <- 12
#iter   <- 1000

mod3 <- stan_surv(
  formula = Surv(age, status) ~ larvalTreat + 
    adultTreat + sex + (1 | cageID),
  data    = eelife,
  basehaz = "ms",
  chains  = chains, 
  cores   = cores, 
  seed    = seed,
  warmup = 15000,
  iter    = 1e6)
# Took about 12 hrs per chain (in parallel) - 13.12 hrs for the model

# Estimates
print(mod3, digits = 2)

#save(mod3, file="../processed/cageID_rand.Rda")
load("../processed/cageID_rand.Rda")

# Does the random term matter? Fixed model
mod3_fixed <- update(mod3, formula. = Surv(
  age, status) ~ larvalTreat + adultTreat + sex) 
# Note: "formula." re-fit named model
# about 6.85 hrs per chain

summary(mod3_fixed)
ps_check(mod3_fixed)
save(mod3_fixed, file="../processed/cageID_fixed.Rda")
load("../processed/cageID_fixed.Rda")

# Calculate 'loo' (integer overflow error!!!)
lcomp3 <- loo_compare(loo(mod3_fixed),loo(mod3))
            
waic_fixed <- waic(mod3_fixed)
loo_mod3 <- loo(mod3)
loo_compare(loo_fixed, loo_mod3)


# Random intercept and random slope model
mod4 <- stan_surv(formula = Surv(age, status) ~ 
    larvalTreat + adultTreat + sex  + 
    (larvalTreat | cageID) + 
    (adultTreat | cageID) + 
    (sex | cageID),
  data    = eelife,
  basehaz = "ms", basehaz_ops = list(df = 10),
  chains  = chains, 
  cores   = cores, 
  seed    = seed,
  warmup = 15000,
  iter    = 5e5)

#save(mod4, file="../processed/cageID_multi_randints.Rda")
load("../processed/cageID_multi_randints.Rda")

print(mod4, digits = 2)
summary(mod4)


# 'loo' all three models 
loo_randCage <- loo(mod4)
loo_compare(loo_fixed, loo_randint, loo_randtrt)

```

