---
title: "Early Experience Factorial Experiment"
author: "Enoch Ng'oma"
date: "7/4/2018"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
---

# Analysis of diet-dependent developmental plasticity in D. melanogaster lifespan

# Function for changing legend size

```{r}
legSize <- function(myPlot, pointSize = 1.5, textSize = 10, spaceLegend = 1.0) {
    myPlot +
        guides(shape = guide_legend(override.aes = list(size = pointSize)),
               color = guide_legend(override.aes = list(size = pointSize))) +
        theme(legend.title = element_text(size = textSize), 
              legend.text  = element_text(size = textSize),
              legend.key.size = unit(spaceLegend, "lines"))
}

```

# Load packages

```{r}
# remotes::install_github("clauswilke/colorblindr")

library(survival)
library(survminer)
library(tidyverse) 
library(forcats)
library(cowplot)
library(gridExtra)
library(colorblindr)
library(visR)
#library(Greg)


#source("color_map.R")
source("/Users/ngomae/MyGithub/Functions_and_Themes/color_map.R")
source("/Users/ngomae/MyGithub/Functions_and_Themes/ggplot_theme.R")

```

## Load data & combine female and male events

```{r}
# Female data
eelife <- read.table('~/MyGithub/early_experience/processed/Female_events_eelife.txt',
                     sep = "\t", header = TRUE,
                     stringsAsFactors = FALSE) 

eelifeF <- select(eelife, -setDate, -flipDate, -age)
names(eelifeF)[1:2]<-c("age", "cageID")
eelifeF$sex <- "F"

# Male data
eelifeM <- read.table('~/MyGithub/early_experience/processed/Male_events_eelife.txt',
                     sep = "\t", header = TRUE,
                     stringsAsFactors = FALSE) 

eelifeM <- select(eelife, -setDate, -flipDate, -age)
names(eelifeM)[1:2]<-c("age", "cageID")
eelifeM$sex <- "M"
  
eelife <- rbind(eelifeF,eelifeM)
  

#eelife$carried <- 0 
eelife <- na.omit(eelife)

eelife$status <- ifelse(eelife[,"status"] == 2, 1, 0)

write.csv(eelife, file="~/MyGithub/early_experience/processed/eeAlldat.csv, row.names = FALSE")
```

# Testing developmental theories

## a) Survival
#### Males and females together
#### Females alone
#### Males alone
## b) Reaction norms


# Male and female combined - Summary

```{r}

eelife <- read.csv("~/MyGithub/early_experience/processed/eeAlldat.csv") %>%
  select(-X)

# Look at distribution of events
eelife <- eelife %>% 
  mutate(stats = ifelse(status == 2, 1, 0)) %>%
  select(cageID,larvalTreat,adultTreat,age,status,sex)

eelife <- unite(eelife, larv_adult,larvalTreat,adultTreat,sep = "_",remove = FALSE)

mp <- survfit(Surv(age, status) ~ larvalTreat + adultTreat,
                   conf.type = "log", 
                   conf.int = 0.95,
                   type = "kaplan-meier",
                   error = "greenwood",
                   data = eelife)

mp
summary(mp)$table[, "median"]

mp.plot <- ggsurvplot(mp, data = eelife, 
                     conf.int = FALSE,
                     break.time.by = 20,
                     palette = c("#999999", "#E69F00", 
                                 "#56B4E9", "#009E73"),
                     fun = "pct", 
                     surv.median.line = "hv",
                     font.x = 12,
                     font.y = 12,
                     legend = c(0.9, 0.6),
                     #risk.table = TRUE,
                     risk.table = "percentage",
                     risk.table.fontsize = 4,
                     tables.height = 0.35,
                     tables.theme = theme_cleantable(),
                     legend.labs = c("HP_HP","HP_LP","LP_HP","LP_LP"),
                     legend.title = ("Diet \n combination")) +
                      xlab("Age (days)")
mp.plot$table <- mp.plot$table + labs(y = NULL)
mp.plot
#mp.plot$table + theme(plot.title = element_text(size=12))

# Combine the main plot and risk table into one grob
combined_plot <- arrangeGrob(mp.plot$plot, mp.plot$table, ncol = 1, heights = c(2, 1))

ggsave("~/MyGithub/early_experience/plots/SurvPlots_for_fullModel_with_RiskTable_All.png",
       plot = combined_plot, width = 8, height = 8, dpi = 300)


# Significance testing - overall and pairwise
survdiff(Surv(age) ~larvalTreat + adultTreat, 
                     data = eelife, rho = 0) 

pairwise_diffs <- pairwise_survdiff(Surv(age) ~ larv_adult, 
                                      data = eelife, 
                                      p.adjust.method = "BH")

pairwise_diffs


summary(mp,times=25)

quantile(mp,  probs = c(0.1, 0.5, 0.9))
visR::get_quantile(mp, probs = c(0.1, 0.5, 0.9))


mps <- surv_fit(Surv(age, status) ~ larvalTreat + adultTreat,
                   data = eelife,group.by="sex")
surv_pvalue(mps)

mpc <- surv_fit(Surv(age, status) ~ sex,
                   data = eelife,group.by="larv_adult")
surv_pvalue(mpc)

# Extract summary table at interesting probabilities
mpsum <- summary(mp,probs=c(0.1,0.5,0.9))
#summary(mp,times=c(1,2,3))


```

# Male and female - by sex

```{r}
mf <- survfit(Surv(age, status) ~ larvalTreat + adultTreat + sex,
                   conf.type = "log", 
                   conf.int = 0.95,
                   type = "kaplan-meier",
                   error = "greenwood",
                   data = eelife)
#summary(mf)
summary(mf)$table[, "median"]
sum_survfit <- surv_summary(mf, data = eelife)

write.csv(sum_survfit, 
          file = "../processed/SummarySurvModel_larvAdulSex.csv",
          row.names = FALSE) 

# Write quantiles
write.csv(quantile(mf, probs = c(0.25, 0.5, 0.75)), 
          file = "../processed/quantiles.csv")

write.csv(quantile(mf, probs = c(0.9, 0.5, 0.1)), 
          file = "../processed/mini-med-max.csv")

# OkabeIto palette in Hex: "#000000","#E69F00","#56B4E9",
# "#009E73", #F0E442","#0072B2","#D55E00","#CC79A7"

mf.plot <- ggsurvplot(mf, data = eelife, 
                     conf.int = FALSE,
                     break.time.by = 20,
                     palette = c("#999999", "#E69F00", "#56B4E9", "#009E73", 
                       "#F0E442", "#0072B2", "#D55E00", "#CC79A7"),
                     fun = "pct", 
                     surv.median.line = "hv",
                     font.x = c(12),
                     font.y = c(12),
                     font.caption = c(10), 
                     font.legend = c(10), 
                     font.tickslab = c(10),
                     legend = c(0.9, 0.6),
                     risk.table = "percentage",
                     #tables.col = "strata",
                     risk.table.y.text = FALSE,
                     risk.table.y.text.col = TRUE,
                     risk.table.fontsize = 4,
                     #tables.y.text = TRUE,
                     #tables.y.fontsize = 3,
                     tables.height = 0.4,
                     tables.theme = theme_cleantable(),
                     legend.labs = c("HP_HP_F","HP_HP_M",
                                     "HP_LP_F","HP_LP_M",
                                     "LP_HP_F","LP_HP_M",
                                     "LP_LP_F","LP_LP_M"),
                     legend.title = "Diet-Sex \n combination") +
                     #guides(colour = guide_legend(nrow = 8)) +
                     xlab("Age (days)")

mf.plot$table <- mf.plot$table + labs(y = NULL)
mf.plot

# mf.plot$table <- mf.plot$table + 
#   theme(
#     axis.text.y = element_blank(),
#     axis.ticks.y = element_blank()
#     )
quantile(mf,  probs = c(0.1, 0.5, 0.9))

ggsave(filename = "~/MyGithub/early_experience/plots/SurvPlots_for_fullModel_with_RiskTable_Sex.png", last_plot(),device = "png")

# Test significance
survdiff(Surv(age) ~larvalTreat+adultTreat+sex,
                     data = eelife, rho = 0) 


# Combine larvalTreat, adultTreat, and sex into a single grouping factor
eelife$group <- interaction(eelife$larvalTreat, eelife$adultTreat, eelife$sex, sep = "_")

# Pairwisediff() using the grouping variable
pairwise_results <- pairwise_survdiff(Surv(age, status) ~ group, 
                                      data = eelife, 
                                      p.adjust.method = "BH")
pairwise_results

# Extract table
library(officer)
library(flextable)

p_values_df <- as.data.frame(as.table(pairwise_results$p.value))
ft <- flextable(p_values_df)
ft <- autofit(ft)
ft

```

# Arrange supplementsary plots

```{r}
# Arrange supplementary figures

suplots <- list()
suplots[[1]] <- mp.plot
suplots[[2]] <- mf.plot

suplots_out <- arrange_ggsurvplots(
  suplots,
  print = FALSE,
  ncol = 2,
  nrow = 1,
  risk.table.height = 0.4)
#ggpar(suplots_out, font.main=10)

ggsave(
  suplots_out,
  file = "~/MyGithub/early_experience/plots/Sup_SurvPlots_for_larvalDiet.pdf",
  width = 14,
  height = 8)
```

# Check if survival is different up to time, t

```{r}
# Truncate time at 20 days and recode events
eelife$time20 <- pmin(eelife$age, 20)
eelife$event20 <- ifelse(eelife$age > 20, 0, eelife$event)

# Perform the survival difference test using the truncated data
survdiff(Surv(time20, event20) ~ larvalTreat + adultTreat, data = eelife, rho = 0)



```

##################
```{r}
mfsum <- summary(mf,probs=c(0.1,0.5,0.9))
cols <- lapply(c(2:7, 10,15:16) , function(x) mfsum[x])
mftab <- do.call(data.frame, cols)
mftab$surv <- round(mftab$surv,digits = 1)
head(mftab)

mfilt <- subset(mftab, surv %in% c(0.9,0.5,0.1))
mfilt$strata <- gsub(",[^=]+=", "_",mfilt$strata) # remove everything between ',' and '=', add '_'
mfilt$Diet <- gsub(" .*","",mfilt$strata) # remove everything before '='
mfilt$Diet <- gsub(".*=", "",mfilt$Diet) # Remove all before and up to "=":



# Aggregate by treatment
#mpfilt$strata <- as.factor(mpfilt$strata)
mfeval <- mfilt %>%
  group_by(Diet,surv) %>%
  summarise_at(c("time", "std.err", "lower", "upper"), mean, na.rm = TRUE)

```
###################

# EarlyRich (C diet) 

### Females only

```{r}
earlyRichF <- subset(eelife, c(eelife$larvalTreat=="C" & eelife$sex=="F"))

erf <- survfit(Surv(age, status) ~ adultTreat,
                   conf.type = "log", 
                   conf.int = 0.95,
                   type = "kaplan-meier",
                   error = "greenwood",
                   data = earlyRichF)

summary(erf)$table[, "median"]
erf_bar <- surv_median(erf)

colnames(erf_bar) <- c("adultDiet", "Median","CI_Lower","CI_upper")
erf_bar$adultDiet <- as.factor(erf_bar$adultDiet)
#rownames(epf_bar) <- c("C", "LP")
levels(erf_bar$adultDiet) <- c("C", "LP")

erfbarp <- ggplot(erf_bar, aes(x=adultDiet, y=Median, fill=adultDiet)) +
  geom_col(color="NA", width=0.5) + 
  # color="black" sets outlines of bars to black
  # color=NA removes lines around bars
  scale_fill_manual(values=c("#999999", "#E69F00")) +
  labs(title="Larval C", x="Adult Diet", y="Median lifespan, days") +
  theme_half_open() +
  my_theme
 
erfbarp

erf.plot <- ggsurvplot(erf, data = earlyRichF, 
                     conf.int = FALSE,
                     break.time.by = 20,
                     palette = c("#999999", "#E69F00", 
                                 "#56B4E9", "#009E73"),
                     fun = "pct", 
                     surv.median.line = "hv",
                     font.x = 12,
                     font.y = 12,
                     font.tickslab = 10,
                     font.main = 10,
                     risk.table = FALSE,
                     legend = c(0.8, 0.8), 
                     title = "Larval diet: C \n (a) Females",
                     legend.labs = c("C", "LP"),
                     legend.title = "Adult Diet") +
  xlab("Age (days)") 
erf.plot

# Test significance
survdiff(Surv(age, status) ~ adultTreat,
                     data = earlyRichF, rho = 0) 

```

# EarlyRich (C diet) 

### Males only

```{r}
earlyRichM <- subset(eelife, c(eelife$larvalTreat=="C" & eelife$sex=="M"))

erm <- survfit(Surv(age, status) ~ adultTreat,
                   conf.type = "log", 
                   conf.int = 0.95,
                   type = "kaplan-meier",
                   error = "greenwood",
                   data = earlyRichM)

summary(erm)
summary(erm)$table[, "median"]
erm_bar <- surv_median(erm)

colnames(erm_bar) <- c("adultDiet", "Median","CI_Lower","CI_upper")
erm_bar$adultDiet <- as.factor(erm_bar$adultDiet)
#rownames(erm_bar) <- c("C", "LP")
levels(erm_bar$adultDiet) <- c("C", "LP")

ermbarp <- ggplot(erm_bar, aes(x=adultDiet, y=Median, fill=adultDiet)) +
  geom_col(color="NA", width=0.5) + 
  # color="black" sets outlines of bars to black
  # color=NA removes lines around bars
  scale_fill_manual(values=c("#999999", "#E69F00")) +
  labs(title="Larval C", x="Adult Diet", y="Median lifespan, days") +
  theme_half_open() +
  my_theme
 
ermbarp 

erm.plot <- ggsurvplot(erm, data = earlyRichM, 
                     conf.int = FALSE,
                     break.time.by = 20,
                     palette = c("#999999", "#E69F00", 
                                 "#56B4E9", "#009E73"),
                     fun = "pct", 
                     #pval = TRUE,
                     surv.median.line = "hv",
                     font.x = 12,
                     font.y = 12,
                     font.tickslab = 10,
                     font.main = 10,
                     risk.table = FALSE,
                     legend="none",
                     # legend = c(0.9, 0.8),                     
                     # legend.labs = c("C", "LP"),
                     legend.title="",
                     title = "Larval diet: C \n (b) Males") +
  xlab("Age (days)") 
erm.plot

survdiff(Surv(age) ~ adultTreat,
                     data = earlyRichM, rho = 0) 

```

# EarlyPoor (LP diet) 

### Females only

```{r}
earlyPoorF <- subset(eelife, c(eelife$larvalTreat=="LP" & eelife$sex=="F"))

epf <- survfit(Surv(age, status) ~ adultTreat,
                   conf.type = "log", 
                   conf.int = 0.95,
                   type = "kaplan-meier",
                   error = "greenwood",
                   data = earlyPoorF,)
summary(epf)$table[, "median"]
epf_bar <- surv_median(epf)

colnames(epf_bar) <- c("adultDiet", "Median","CI_Lower","CI_upper")
epf_bar$adultDiet <- as.factor(epf_bar$adultDiet)
#rownames(epf_bar) <- c("C", "LP")
levels(epf_bar$adultDiet) <- c("C", "LP")

epfbarp <- ggplot(epf_bar, aes(x=adultDiet, y=Median, fill=adultDiet)) +
  geom_col(color="NA", width=0.5) + 
  # color="black" sets outlines of bars to black
  # color=NA removes lines around bars
  scale_fill_manual(values=c("#999999", "#E69F00")) +
  labs(title="Larval LP", x="Adult Diet", y="Median lifespan, days") +
  theme_half_open() +
  my_theme
 
epfbarp 

epf.plot <- ggsurvplot(epf, data = earlyPoorF, 
                     conf.int = FALSE,
                     break.time.by = 20,
                     palette = c("#999999", "#E69F00", 
                                 "#56B4E9", "#009E73"),
                     fun = "pct", 
                     #pval = TRUE,
                     surv.median.line = "hv",
                     font.x = 12,
                     font.y = 12,
                     font.tickslab = 10,
                     font.main = 10,
                     risk.table = FALSE,
                     legend="none",
                     # legend = c(0.9, 0.8),                     
                     # legend.labs = c("C", "LP"),
                     # legend.title = "Adult Diet",
                     title = "Larval diet: LP \n (c) Females") +
  xlab("Age (days)") +
  ylab("")
epf.plot

# Test significance
survdiff(Surv(age, status) ~ adultTreat,
                     data = earlyPoorF, rho = 0) 
summary(epf)
summary(mf)$table[, "median"]
```

### Males only

```{r}
earlyPoorM <- subset(eelife, c(eelife$larvalTreat=="LP" & eelife$sex=="M"))

epm <- survfit(Surv(age, status) ~ adultTreat,
                   conf.type = "log", 
                   conf.int = 0.95,
                   type = "kaplan-meier",
                   error = "greenwood",
                   data = earlyPoorM)

summary(epm)$table[, "median"]
epm_bar <- surv_median(epm)

colnames(epm_bar) <- c("adultDiet", "Median","CI_Lower","CI_upper")
epm_bar$adultDiet <- as.factor(epm_bar$adultDiet)
#rownames(epm_bar) <- c("C", "LP")
levels(epm_bar$adultDiet) <- c("C", "LP")

epmbarp <- ggplot(epm_bar, aes(x=adultDiet, y=Median, fill=adultDiet)) +
  geom_col(color="NA", width=0.5) + 
  # color="black" sets outlines of bars to black
  # color=NA removes lines around bars
  scale_fill_manual(values=c("#999999", "#E69F00")) +
  labs(title="Larval LP", x="Adult Diet", y="Median lifespan, days") +
  theme_half_open() +
  my_theme
 
epmbarp

epm.plot <- ggsurvplot(epm, data = earlyPoorM, 
                     conf.int = FALSE,
                     break.time.by = 20,
                     palette = c("#999999", "#E69F00", 
                                 "#56B4E9", "#009E73"),
                     fun = "pct", 
                     #pval = TRUE,
                     surv.median.line = "hv",
                     font.x = 12,
                     font.y = 12,
                     font.tickslab = 10,
                     font.main = 10,
                     risk.table = FALSE,
                     legend="none",
                     #risk.table.pos = "in",      
                     # ggtheme = theme_classic2(base_family = "Courier New"),
                     # font.family = "Courier New",
                     # legend = c(0.9, 0.8),                     
                     # legend.labs = c("C", "LP"),
                     # legend.title = "Adult Diet",
                     title = "Larval diet: LP \n (d) Males") +
  xlab("Age (days)") +
  ylab("")
epm.plot

survdiff(Surv(age, status) ~ adultTreat,
                     data = earlyPoorM, rho = 0) 
```

# Arrange survival plots

```{r}
splots <- list()
splots[[1]] <- erf.plot
splots[[2]] <- erm.plot
splots[[3]] <- epf.plot
splots[[4]] <- epm.plot

splots_out <- arrange_ggsurvplots(
  splots,
  print = FALSE,
  ncol = 2,
  nrow = 2)
#ggpar(splots_out, font.main=10)

ggsave(
  splots_out,
  file = "~/MyGithub/early_experience/plots/SurvPlots_for_larvalDiet.pdf",
  width = 5,
  height = 4)

```

# Survival at 90%, 50% and 10% survival

```{r}
q.erf <- visR::get_quantile(erf, probs = c(0.1, 0.5, 0.9))
q.erf$sex <- "F"
q.erm <- visR::get_quantile(erm, probs = c(0.1, 0.5, 0.9))
q.erm$sex <- "M"
q.epf <- visR::get_quantile(epf, probs = c(0.1, 0.5, 0.9))
q.epf$sex <- "F"
q.epm <- visR::get_quantile(epm, probs = c(0.1, 0.5, 0.9))
q.epm$sex <- "M"
quants <- rbind(q.erf,q.erm,q.epf,q.epm)

q.mf <- visR::get_quantile(mf, probs = c(0.1, 0.5, 0.9))


q.mf <- q.mf %>%
  rename(Diet = strata,CI = quantity)

q.mf$Diet <- gsub(",[^=]+=", "_",q.mf$Diet) 
q.mf$Diet <- gsub(".*=","",q.mf$Diet)
q.mf$CI <- gsub("^(\\w).*$","\\1",q.mf$CI)

# q.mf$Diet <- gsub(".*=","",q.mf$Diet)
# q.mf$CI <- gsub("^(\\w).*$","\\1",q.mf$CI)

gathercols <- c("10", "50", "90")

q_long <- q.mf %>%
  gather(quantile, surv_prob, gathercols) %>%
  filter(CI=="q") 
q_long$sex <- str_split(q_long$Diet, "_", simplify = TRUE)[,3]
#q_long$quantile <- as.numeric(q_long$quantile)

qx <- ggplot(q_long, aes(x=surv_prob, y=quantile, group_by=Diet, shape=sex,color=Diet)) + 
  geom_jitter(alpha=0.6,size=4) +
  #facet_grid(rows=vars(quantile), cols=vars(sex), scales= "free_y") +
  #scale_x_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 10)) +
  #scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 25)) +
  scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9", "#009E73", 
                       "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) +
  labs(x = "No. of days", y = "Percent survival point") +
  theme(legend.position = "none",
        panel.border = element_rect(color="black,size=1")) +
  #theme(panel.border = element_rect(color="black,size=1")) +
  my_theme +
  theme_half_open()
al <- legSize(qx)
al 

ggsave(
  al,
  file = "~/MyGithub/early_experience/plots/SurvPlots_at_3quants.pdf",
  width = 5,
  height = 4)

# Females only
qxf <- subset(q_long, q_long$sex=="F")
qf <- ggplot(qxf, aes(x=surv_prob, y=quantile, group_by=Diet, color=Diet)) + 
  geom_jitter(alpha=0.6,size=4) +
  #facet_grid(rows=vars(quantile), cols=vars(sex), scales= "free_y") +
  #scale_x_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 10)) +
  #scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 25)) +
  scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9", "#009E73", 
                       "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) +
  labs(x = "No. of days", y = "Percent survival point") +
  theme(legend.position = "none",
        panel.border = element_rect(color="black,size=1")) +
  #theme(panel.border = element_rect(color="black,size=1")) +
  my_theme +
  theme_half_open()
af <- legSize(qf)
af 

ggsave(
  af,
  file = "~/MyGithub/early_experience/plots/SurvPlots_at_3quants_female.pdf",
  width = 4,
  height = 3)

# Males only
qxm <- subset(q_long, q_long$sex=="M")
qm <- ggplot(qxm, aes(x=surv_prob, y=quantile, group_by=Diet, color=Diet)) + 
  geom_jitter(alpha=0.6,size=4) +
  #facet_grid(rows=vars(quantile), cols=vars(sex), scales= "free_y") +
  #scale_x_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 10)) +
  #scale_y_continuous(limits = c(0, 100), breaks = seq(0, 100, by = 25)) +
  scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9", "#009E73", 
                       "#F0E442", "#0072B2", "#D55E00", "#CC79A7")) +
  labs(x = "No. of days", y = "Percent survival point") +
  theme(legend.position = "none",
        panel.border = element_rect(color="black,size=1")) +
  #theme(panel.border = element_rect(color="black,size=1")) +
  my_theme +
  theme_half_open()
am <- legSize(qm)
am 

ggsave(
  am,
  file = "~/MyGithub/early_experience/plots/SurvPlots_at_3quants_male.pdf",
  width = 4,
  height = 3)
```




