---
title: "Lifetime fecundity"
author: "Enoch Ng'oma"
date: "March 10, 2020"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}

library(tidyverse) 
library(cowplot)
library(gplots)
library(gdata)
library(gtools)
library(MuMIn)

source("/Users/ngomae/MyGithub/Functions_and_Themes/ggplot_theme.R")
```

# Fecundity data

```{r}
eggs <- read.table("~/MyGithub/early_experience/processed/tricount.txt",
            sep = '\t', header = TRUE)

zdn <-eggs %>%  
  dplyr::select(id,age,cam_id,larvalTreat,adultTreat,
                NstartF,alive,aj1_count,aj2_count,dd1_count,numEggs,
                eggpDay,eggpFemDay)
head(zdn)

zdn %>% group_by(id) %>% summarise(sum(numEggs),n=n())
hist(log(zdn$numEggs))

```

# Age-specific fecundity

```{r}

ltime <- zdn %>% unite("larad", larvalTreat, 
               adultTreat, sep = "_", remove=FALSE) %>%
  select(id, larad, age,  larvalTreat, adultTreat, 
                numEggs, eggpDay, eggpFemDay) %>%
  rename(cage="id")

ltime$cage <- as.factor(ltime$cage)
ltime$larad <- as.factor(ltime$larad)

fii <- ltime %>% group_by(larad) %>%
  summarise(Samu = sum(numEggs))
fii

pairwise.t.test(ltime$numEggs, ltime$larad)

```

# Effect of larval treatment on fecundity

- Larval diet has effect (***, ncq)
- Some effect of cage (*, ncq2)
- DR_C combination increases fecundity (***,nqc3b)
- Larval DR has effect(**,nql1,nql3)
- Adult DR has some effect (*. nql2)
- Larval DR interacts with adult DR (**,nql4)
- Interaction disappears if adult diet included as main effect (nql5)

```{r}
nqc1 <- aov(eggpFemDay ~ larad, data = ltime)
nqc2 <- aov(eggpFemDay ~ larad + cage, data = ltime)

plotmeans(eggpFemDay ~ cage, data = ltime)
plotmeans(eggpFemDay ~ larad, data = ltime)

# Effect of larval diet
nql1 <- lm(eggpFemDay ~ larvalTreat, data = ltime)
nql2 <- lm(eggpFemDay ~ adultTreat, data = ltime)
nql3 <- lm(eggpFemDay ~ larvalTreat + adultTreat, data = ltime)
nql4 <- lm(eggpFemDay ~larvalTreat + larvalTreat:adultTreat, data = ltime)
nql5 <- lm(eggpFemDay ~larvalTreat + adultTreat + 
             larvalTreat:adultTreat, data = ltime)

summary(nql1)
summary(nql2)
summary(nql3)
summary(nql4)
summary(nql5)

```

# Dredge a saturated per female per 3-hr model

```{r}
# Per female per day
egg_lm <- lm(eggpFemDay ~ larad*age + larad*cage, data=ltime)

# Per female per day - dredge

options(na.action = "na.fail") #required for dredge to run
egg_dr <- dredge(egg_lm, evaluate=TRUE,rank=AICc,trace=1)
options(na.action = "na.omit") #set back to default
nrow(egg_dr)
head(egg_dr)

top_egg_dr <- get.models(egg_dr, subset=1)[[1]]
top_egg_dr
#head(top_egg_dr)
summary(top_egg_dr)

avmod <- summary(model.avg(egg_dr, subset = delta <= 2))
```

# Dredge a saturated age-specific fecundity

```{r}
egg_lft <- lm(numEggs ~ larad*age + larad*cage, data=ltime)

options(na.action = "na.fail") #required for dredge to run
lft_dr <- dredge(egg_lft, evaluate=TRUE,rank=AICc,trace=1)
options(na.action = "na.omit") #set back to default
nrow(lft_dr)
head(lft_dr)

top_lft_dr <- get.models(lft_dr, subset=1)[[1]]
top_lft_dr
#head(top_lft_dr)
summary(top_lft_dr)

avmod_lft <- summary(model.avg(lft_dr, subset = delta <= 2))
```

