---
title: "GLMM"
author: "Enoch Ng'oma"
date: "8/6/2019"
output: 
  html_document:
    toc: true
    toc_depth: 2 #defaults to 3 if not specified
    toc_float: true
    number_sections: true
    fig_width: 7
    fig_height: 6
    fig_caption: true
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(car)
library(MASS)
library(lme4)
library(tidyverse)
library(cowplot)
library(MCMCglmm)

source("color_map.R")

set.seed(23025)

```

# Data for males and females together

```{r}

eelife <- read.csv("../processed/eeAlldat.csv", header=TRUE)
names(eelife)[1]<-"event"
head(eelife)
```

# Density plots

```{r}
dev.off()
ggplot(eelife, aes(x = age.a)) + 
  geom_density() + 
  theme_half_open() +
  facet_wrap(larv_adult ~ cageID)
```

# Distribution fitting age data
# Ref.: https://ase.tufts.edu/gsc/gradresources/guidetomixedmodelsinr/mixed%20model%20guide.html

```{r}
eelife$age.a <- eelife$age + 1
qqp(eelife$age.a, "weibull")

qqp(eelife$age.a, "lnorm")

nbinom <- fitdistr(eelife$age.a, "Negative Binomial")
qqp(eelife$age, "nbinom", size = nbinom$estimate[[1]], mu = nbinom$estimate[[2]])

poissonn <- fitdistr(eelife$age.a, "Poisson")
qqp(eelife$age.a, "pois", lambda=1,poissonn$estimate)

gammaa <- fitdistr(eelife$age.a, "gamma")
qqp(eelife$age.a, "gamma", shape = gammaa$estimate[[1]], rate = gammaa$estimate[[2]])

```

# Test penalized quasilikelihood (PQL)?

# PQL is a flexible technique that can deal with non-normal data, unbalanced design, and crossed random effects. However, it produces biased estimates if your response variable fits a discrete count distribution, like Poisson or binomial, and the mean is less than 5 - or if your response variable is binary

```{r}
pql1 <- glmmPQL(age.a ~ larv_adult, random =  ~1 | cageID, 
               family = gaussian(link = "log"),
               data = eelife, verbose = FALSE)
summary(pql1)

```

Fixed effects: age.a ~ larv_adult 
                   Value  Std.Error   DF   t-value p-value
(Intercept)     3.550191 0.03431733 3124 103.45183  0.0000
larv_adultC_DR  0.145601 0.04716790    4   3.08687  0.0367
larv_adultDR_C  0.051503 0.04695323    4   1.09691  0.3343
larv_adultDR_DR 0.149532 0.04684138    4   3.19232  0.0331

Diet swith from C in larvae to DR in adults seems to extend lifespan
Maintaining DR diet in both stages seems to extend lifespan
However, a switch from DR to C seems to have no effect on lifespan

```{r}
pql2 <- glmmPQL(age.a ~ larv_adult + sex, ~1 | cageID, 
               family = gaussian(link = "log"),
               data = eelife, verbose = FALSE)

```

Fixed effects: age.a ~ larv_adult + sex 
                   Value  Std.Error   DF  t-value p-value
(Intercept)     3.488111 0.03680965 3123 94.76077  0.0000
larv_adultC_DR  0.129301 0.04951595    4  2.61129  0.0593
larv_adultDR_C  0.055970 0.04930424    4  1.13520  0.3197
larv_adultDR_DR 0.151264 0.04916373    4  3.07675  0.0370
sexM            0.159810 0.01837741 3123  8.69598  0.0000

A large effect of sex! The male xes appears to affect lifespan positively - longer in males
The effect of diet switch from C in larvae to DR in adults on lifespan is marginal in the presence of sex
Maintaining DR diet in both stages seems to extend lifespan even in the presence of sex
A switch from DR to C remains of no effect on lifespan

```{r}
pql3 <- glmmPQL(age.a ~ larv_adult + sex, ~1 | cageID/adultTreat, 
               family = gaussian(link = "log"),
               data = eelife, verbose = FALSE)

```

Nesting larvalTreat or adultTreat in cageID yields identical effects.

```{r}
set.seed(37264)

iter <- 2e5
burnin <- 15000
rerun_MCMC <- FALSE

prior = list(R = list(V = 1, n = 0, fix = 1), G = list(G1 = list(V = 1, n = 1),
    G2 = list(V = 1, n = 1), G3 = list(V = 1, n = 1)))

eemcmc <- MCMCglmm(age.a ~ 1, 
                   random = ~larv_adult + cageID + sex,
                   data = eelife, 
                   family = "gaussian", 
                   prior = prior, 
                   verbose = FALSE)

summary(eemcmc)
```

A huge mean for sex (68.2), but some effect also of cage (mean = 9.902) and diet switch (4.669). The probability of living to a given age varies a lot between sex, and also between cages and diet switch.



# Females only analysis

```{r}
eelifeF <- subset(eelife, eelife$sex=="F")
```

```{r}
fpl1 <- glmmPQL(age ~ larv_adult, random =  ~1 | cageID, 
               family = gaussian(link = "log"),
               data = eelifeF, verbose = FALSE)
summary(fpl1)

```

No effect of diet in all combinations

```{r}
fpl2 <- glmmPQL(age ~ larv_adult, ~1 | cageID / larvalTreat, 
               family = gaussian(link = "log"),
               data = eelifeF, verbose = FALSE)

```

No diet effects

```{r}
fpl3 <- glmmPQL(age ~ larv_adult, ~1 | cageID/adultTreat, 
               family = gaussian(link = "log"),
               data = eelifeF, verbose = FALSE)

```

No diet effects

```{r}
set.seed(27014)

iter <- 2e5
burnin <- 15000
rerun_MCMC <- FALSE

prior = list(R = list(V = 1, n = 0, fix = 1), 
             G = list(G1 = list(V = 1, n = 1), 
                      G2 = list(V = 1, n = 1),
                      G3 = list(V = 1, n = 1),
                      G4 = list(V = 1, n = 1)))

eemcmc <- MCMCglmm(age ~ 1, 
                   random = ~larv_adult + cageID,+ larvalTreat + adultTreat,
                   data = eelifeF, 
                   family = "gaussian", 
                   prior = prior, 
                   verbose = FALSE)

summary(eemcmc)
```
