---
title: "Mixed effects Cox Regression"
author: "Enoch Ng'oma"
date: "7/4/2018"
output:
  html_document:
    theme: flatly
    toc: true
    toc_float:
      collapsed: false
      smooth_scroll: true
editor_options: 
  chunk_output_type: console
---

# Analysis of diet-dependent developmental plasticity in D. melanogaster lifespan

# Load packages

```{r}
library(survival)
#library(coxme)
library(rstanarm)
library(brms)
library(tidyverse)
library(cowplot)
library(tidybayes)
library(bayesplot)
library(broom)

#source("color_map.R")
#source("ggplot_theme.R")
```

## Load data & combine female and male events

```{r}

# Data has both male and female events

eelife <- read.csv("../processed/eeAlldat.csv")

```

# Basic model

age ~ larv + adult +sex + cluster(larv) : LarvalTreatDR becomes very sig

```{r}

M <- coxph(Surv(age, status==2) ~ larvalTreat + adultTreat + sex, 
                   data = eelife)
M

# Model with robust SE via clustering
M2 <- coxph(Surv(age, status==2) ~ larvalTreat + 
              adultTreat + sex + cluster(larvalTreat), data = eelife)
M2
``` 

# Mixed effects model

```{r}
M3 <- coxme(Surv(age, status==2) ~ larvalTreat + adultTreat + 
              sex  + (1 | cageID), data = eelife)
M3

M4 <- coxme(Surv(age, status==2) ~ larvalTreat + sex  + 
              (1 | adultTreat), data = eelife)
M4

M5 <- coxme(Surv(age, status==2) ~ larvalTreat + adultTreat + sex +
              (1 | cageID), data = eelife)
M5

M6 <- coxme(Surv(age, status==2) ~ larvalTreat + adultTreat +
              (1 | sex), data = eelife)
M6

M7 <- coxme(Surv(age, status==2) ~ larvalTreat + adultTreat +
              (1 | sex) + (1 | cageID), data = eelife)
M7
```


# Bayesian survival models

# rstanarm

## Non-PHs modelled with a smooth cubic B-spline function for the time-varying log hazard ratio

```{r}
set.seed(989898)
CHAINS <- 1
CORES  <- 1
SEED   <- 12345
ITER   <- 1000

# Baseline hazard function - assuming covariate effect of zero
mod0 <- stan_surv(formula = Surv(age, status==2) ~ tve(larvalTreat) + 
                     tve(adultTreat) + tve(sex), 
                  data    = eelife,
                  chains  = CHAINS, 
                  cores   = CORES, 
                  seed    = SEED,
                  iter    = ITER)
print(mod0, digits = 3)
plot(mod0) + coord_cartesian(ylim = c(0, 0.1))


# Test different parametric hazards
# i.e. fit several models, each with a different baseline hazard

mod1_exp      <- update(mod0, basehaz = "exp") #47min
mod1_weibull  <- update(mod0, basehaz = "weibull") # ~ 2hrs
mod1_gompertz <- update(mod0, basehaz = "gompertz") # 94s - not converged
mod1_bspline  <- update(mod0, basehaz = "bs") # 98.5 min
mod1_mspline1 <- update(mod0, basehaz = "ms") # 52.9 min
mod1_mspline2 <- update(mod0, basehaz = "ms", 
                        basehaz_ops = list(df = 10))

# save(mod0,mod1_exp,mod1_weibull,mod1_gompertz,mod1_bspline,
#        mod1_mspline1,mod1_mspline2, file="../processed/BaseHaz.Rda")


# Plot the baseline hazards with 95% posterior uncertainty limits
load("../processed/BaseHaz.Rda")

plotfun <- function(model, title) {
  plot(model, plotfun = "basehaz") +              
    coord_cartesian(ylim = c(0,0.4)) +            
    labs(title = title) +                        
    theme(plot.title = element_text(hjust = 0.5,size = 11),
          axis.title = element_text(size = 10)) + 
    theme_set(theme_cowplot()) +
    theme_half_open() 
}
p_exp      <- plotfun(mod1_exp,      title = "Exponential")
p_weibull  <- plotfun(mod1_weibull,  title = "Weibull")
p_gompertz <- plotfun(mod1_gompertz, title = "Gompertz")
p_bspline  <- plotfun(mod1_bspline,  title = "B-splines with df = 5")
p_mspline1 <- plotfun(mod1_mspline1, title = "M-splines with df = 5")
p_mspline2 <- plotfun(mod1_mspline2, title = "M-splines with df = 10")
cs <-bayesplot::bayesplot_grid(p_exp,
                          p_weibull,
                          p_gompertz,
                          p_bspline,
                          p_mspline1,
                          p_mspline2,
                          grid_args = list(ncol = 3))

ggplot2::ggsave(cs,filename="../plots/baseHaz_tve_fixedEffs.pdf",height = 6, width=10)

# Compare fits using `loo` for `stansurv` objects

lcomp <- loo_compare(loo(mod1_exp),
            loo(mod1_weibull),
            loo(mod1_gompertz),
            loo(mod1_bspline),
            loo(mod1_mspline1),
            loo(mod1_mspline2))

## Record execution time and multicore use
end_time <- Sys.time()

diff_time <- difftime(end_time, start_time, units = "auto")
cat("Started  ", as.character(start_time), "\n",
    "Finished ", as.character(end_time), "\n",
    "Time difference of ", diff_time, " ", attr(diff_time, "units"), "\n",
    "Used ", foreach::getDoParWorkers(), " cores\n",
    "Used ", foreach::getDoParName(), " as backend\n",
    sep = "")

```

# Non-PHs modelled with a piecewise constant function

```{r}
# Piecewise constant: alternative to model tve() log hazard ratio

mod1 <- stan_surv(formula = Surv(age, status==2) ~ 
                    tve(larvalTreat, degree = 0, knots = 2.5) + 
                    tve(adultTreat, degree = 0, knots = 2.5) + 
                    tve(sex, degree = 0, knots = 2.5),
                    data    = eelife, 
                    chains  = CHAINS, 
                    cores   = CORES, 
                    seed    = SEED,
                    iter    = ITER)

```


# Multilevel models

```{r}

```

